### 2021-04-09 Submission template.Rmd
Submission template

---
header-includes:
   - \usepackage{bm}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage{booktabs}
   - \usepackage{sectsty} \sectionfont{\centering \emph}
output:
  pdf_document: default
  html_document: default
  word_document: default

---
\newcommand\btheta{\mbox{\boldmath${\theta}$}}

```{r setup, include=FALSE}
# Use echo = FALSE for Answer Key without R code, 
# echo = TRUE for complete solutions.
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library("tidyverse")
library("xtable")
library("Rcpp")
library("plyr")
library("knitr")
```

Team No: 202

Track No.: 2.

# Our name of the project

Link to GitHub Repository: https://github.com/MantautasRimkus/StanfordDatathonCSU

### Introduction