### 2021-04-11 Population_Study.Rmd

---
title: "Population Study"
author: "Paul Chong"
date: "4/10/2021"
output: pdf_document
---

\colorlet{shadecolor}{gray!10}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\newcommand{\hstart}{ \colorlet{shadecolor}{orange!20}
\begin{shaded} }
\newcommand{\hstop}{  \end{shaded} \colorlet{shadecolor}{gray!10}}

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
```

```{r}
college_pop <- read.csv("~/Datathon 2021/college_population_data.csv")
total_pop <- read.csv("~/Datathon 2021/18-29_population_data.csv")
college_pop
```

```{r, fig.cap = "The first line is the proportion of students who are between ages 18-29 (blue). The second line is the proportion of 18-29 year olds who are attending a university (red)."}

p <- ggplot(college_pop, aes(x = Year)) +
  geom_line(aes(y = college_pop$Porportion.Inside, color = "blue")) +
  geom_point(aes(y = college_pop$Porportion.Inside, color = "blue"))+
  geom_line(data = total_pop, aes(y = total_pop$Proportion, color = "red")) +
  geom_point(data = total_pop, aes(y = total_pop$Proportion, color = "red"))+
  scale_y_continuous(breaks = c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.00), 
                     limits = c(0,1)) +
  labs(x = "Year", y = "Proportions",  
       title = "Different Measurements for University Students Aged 18-29",
       colour = "Age Groups") + 
  scale_color_manual(labels = c("Students in 18-29", "18-29 in Students"), values = c("blue", "red")) +
  theme_bw()

p
```

\hstart

University students have been found to experience mental health disorders at a disproportionate rate.  One meta-analysis found the average rate of depression in university students to be 30.6\%, when the general US population is estimated to have a 9\% average rate (REF).  The age range of most university students, 18-22, is host to large social and mental developments and is one of the more common time periods for mental disorders like anxiety to manifest (REF).  Adding the pressures of school and high-impact life choices, it is understandable that university students experience more than their fair share of mental health issues.

The year 2020 and the disease Covid-19 added further difficulty to the lives of those who are college aged.  Students who returned or were admitted to universities in fall 2020 were forced to learn new styles of learning and make sacrifices in their social lives.   Universities transitioned to partially or fully online curriculum, adding distance between student and staff communities.  Unsurprisingly, measures for anxiety and depression have increased under the duration of the Covid-19 pandemic.  This report examines the survey-reported changes in anxiety or depression prevalence for college-aged persons in the United States.  Data on Covid-19 infection rates are linked with the mental health surveys.

We decided to focus on individuals between 18-29 because this is the age group that primarily makes up the university population. Although only 30% of individuals aged between 18-29 attend universities, more than 70% of students attending universities in the United States are between 18-29. This has been a consistent trend for the past 10 years, so we can assume that the figures for 2020 is going to be approximately the same. Our team believes that because of the population proportions that we have found for university students, the following findings will apply to most of the students enrolled in universities.

\hstop

\hstart

EXPLANATION ON DATA

The first dataset we used to find university populations was a data frame we found from the National Center for Education Statistics. The dataset records total fall enrollment in degree-granting institutions by race, gender, and age of student. We extracted data from this table to create our own dataframe that contains different age populations for enrollment for every year. Each row (observation) contains the year of the semester, age groups for 14-17, 18-19, 20-21, 22-24, 25-29, 30-34, and 35 and older. Each row also contains the total enrollment for that semester, the proportion that are between ages 18-29, and the proportion that is not in that age range.

The next dataset we utilized for our study was from the United States Census Bureau Household Pulse Survey. This experimental survey was conducted to quickly and efficiently collect data on how people's lives were impacted by COVID-19 pandemic. The survey was conducted over a period of 26 weeks, where in each week there were detailed tables provided of various topics, such as education, employment, health, etc. It should be noted that the term "weeks" is not defined as a 7-day period, but rather a set period of time determined by the researchers. For example, the table for week 1 contains data for April 23 to May 5.

We focused on a table that tracked symptoms of anxiety by different characteristics, such as age, gender, marital status, education levels, etc. Our team extracted the anxiety tables for the age group 18-29 from all 26 tables for every state for each week. We then created a new data frame that contains the anxiety levels for those aged 18-29 for 26 weeks. Each individual was asked how frequent they were feeling nervous, anxious, or on the edge. They were also asked frequency of not being able to stop or control worrying. In each of these questions, the options were "Not at all", "Several days", "More than half the days", "Nearly everyday", or "Did not report".

\hstop